\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{fancyhdr}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage[multiple]{footmisc}

\usepackage{listings}
\lstset{language=Python,
        basicstyle=\footnotesize\ttfamily,
        showspaces=false,
        showstringspaces=false,
        tabsize=2,
        breaklines=false,
        breakatwhitespace=true,
        identifierstyle=\ttfamily,
        keywordstyle=\bfseries,
        commentstyle=\it,
        stringstyle=\it,
    }

\usepackage[pdftex]{graphicx}

% header
\fancyhead{}
\fancyfoot{}
\fancyfoot[C]{\thepage}
\fancyhead[R]{Daniel Foreman-Mackey}
\fancyhead[L]{Biophysics --- Midterm}
\pagestyle{fancy}
\setlength{\headsep}{25pt}

% shortcuts
\newcommand{\Eq}[1]{Equation (\ref{eq:#1})}
\newcommand{\eq}[1]{Equation (\ref{eq:#1})}
\newcommand{\eqlabel}[1]{\label{eq:#1}}
\newcommand{\Fig}[1]{Figure \ref{fig:#1}}
\newcommand{\fig}[1]{Figure \ref{fig:#1}}
\newcommand{\figlabel}[1]{\label{fig:#1}}

\newcommand{\etal}{\emph{et al.}}

% commands
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\unit}[1]{\ensuremath{\,\mathrm{#1}}}
\def\baselinestretch{1.2}

\begin{document}

\section*{Part 1}

Instead of researching the spontaneous organization of biological molecules
like we discussed previously\footnote{I did however look into the research by
N. Clark \emph{et al.} and it's very interesting, thanks!}, after we discussed
it in class, I decided to look into {\bf the structure of chromatin}.
In particular, I was especially interested in the techniques behind
{\bf chromosome conformation capture} (3C) so I'll focus on that.
The other sub-area that I'll consider is {\bf fluorescence \emph{in-situ}
hybridization} (FISH) studies of ``chromosome territories''.

\subsection*{Part 1.3}

\subsubsection*{Chromosome territories}

\emph{Reviews: \cite{ct-review3}, \cite{ct-review2} and \cite{ct-review1}}

\noindent\emph{Key papers: \cite{hamster}, \cite{cremer}, and \cite{local}}

It was long speculated---starting as early as the late 1800s
\cite{old}---that chromosomes are spatially isolated---in regions now called
``chromosome territories''---even during interphase \cite{ct-review1,
ct-review2}.
This theory was at odds with the competing---and then more popular
theory---that chromatin were arranged as a fully-mixed equilibrium
``spaghetti'' in the nucleus and it was mostly disregarded until the 1970s
\cite{cremer}.
The first convincing evidence of the territorial chromosome theory appeared
when members of the Cremer lab in Heidelberg, Germany determined that only a
small number of chromosomes were affected by focused radiation damage applied
to the DNA during interphase \cite{zorn, hamster}.
In the 80s and 90s, this theory was confirmed and more directly studied using
fluorescence \emph{in-situ} hybridization (FISH) \cite{cremer}.
This technique lets researchers tag and image the chromatin of specific
chromosomes, by targeting and linking fluorescent probes to specific DNA
sequences.
This image can then be used to directly see the location of each chromosome
even during interphase.
The fact that chromosomes are localized during interphase certainly has
epigenetic effects and recently, there has been research aimed at determining
and understanding the locations of specific chromosomes \cite{local}.

Much of the research in this field came from brothers {\bf Thomas Cremer}
(University of Munich, Germany) and {\bf Christoph Cremer} (University of
Heidelberg, Germany).
These researchers presented the first compelling evidence for the chromosome
territory theory \cite{hamster}.
In this paper, they subjected cells in interphase to focused UV damage and
found that if they followed the cells through to mitosis, the damage was only
visible on a few specific chromosomes.
In contrast, if the chromosomes get distributed and mixed during interphase
then you would expect the damage to be apparent on all the chromosomes.
This paper has 205 citations%
\footnote{\url{http://scholar.google.com/scholar?cites=1373788124891469070}}.


The Cremer brothers went on to use the FISH method for imaging the cell during
interphase \cite{cremer}.
By carefully tagging specific DNA sequences from different chromosomes with
different colors, this research confirmed the existence of chromosome
territories.
This paper has 490 citations%
\footnote{\url{http://scholar.google.com/scholar?cites=2575496365450001910}}.

Another important researcher in this field is {\bf Wendy Bickmore} (University
of Edinburgh).
A key paper out of her lab \cite{local} demonstrated that the locations of
specific chromosomes in human cells are set early and reproducibly.
In particular, they showed that two similar (in terms of size and the amount
of genetic information content) chromosomes (18 and 19) occupy very different
places from each other but that every cell this same distinction.
This motivates the epigenetic theory that the layout of chromosomes in the
nucleus is actually functional and not random.
This paper has 654 citations%
\footnote{\url{http://scholar.google.com/scholar?cites=4847129722710273741}}.


\subsubsection*{Chromosome conformation capture}

\emph{Reviews: \cite{3c-review1}, \cite{3c-review2}, and \cite{3c-review3}}

\noindent\emph{Key papers: \cite{ccc}, \cite{looping}, \cite{4c}, and
\cite{5c}}

Chromosome conformation capture (3C) is an impressive statistical technique
for determining the distribution of chromatin in the nucleus.
The basic idea is to cross-link chromatin filaments that physically touch and
then sequence the cross-linked strands to determine which parts of the genome
are co-spatial \cite{3c-review1}.
The method is statistical because, in general, you must study a large set of
cells so that random interactions don't dominate the signal \cite{3c-review2}.
The formal steps in the procedure are \cite{3c-review1, 3c-review2}:
\begin{enumerate}

\item Fix the cells with formaldehyde, covalently linking interacting strands
of chromatin,

\item Digest the chromatin with a specific ``restriction enzyme'' that
separates the cross-linked chromatin from the rest,

\item Ligate the cross-linked strands to join them and get a single strand,

\item Dissolve the cross-links and sequence the strands.

\end{enumerate}
After this procedure, the frequency with which sequences are paired is related
to their average physical separation in the nucleus.
The original version of this procedure was published in a seminal 2002 paper
\cite{ccc} where the authors measured the layout of the genome of yeast cells
at various stages of meiosis.
Using this statistical sample of sequence distances, these authors
reconstructed the ensemble-averaged, three-dimensional structure of one of the
chromosomes in the yeast they were studying.

The three key researchers in this field that I would nominate for the Nobel
are {\bf Job Dekker}, {\bf Nancy Kleckner}, and {\bf Wouter de Laat}.
Dekker, now a professor at the University of Massachusetts, developed the 3C
method while a Postdoctoral Fellow in the Kleckner lab at Harvard \cite{ccc}.
Their paper \cite{ccc} described the 3C procedure and, as described above,
they applied it to determine the three-dimensional structure of a chromosome
in yeast.
This is the seminal paper of this field and it has been cited about 1250
times%
\footnote{\url{http://scholar.google.com/scholar?cites=14717049323260909894}}.

The next key paper \cite{looping} is from the de Laat lab.
In this paper, the authors used 3C to study a physical mechanism for
transcription regulation.
In particular, they showed that in liver cells of the murine---where the
$\beta$-globin protein is expressed---the ``locus control region'' (LCR; the
genes responsible for regulating the expression of the $\beta$-globin gene)
comes into physical contact with the $\beta$-globin gene even though it is
located tens of kilobases away in the genome.
In contrast, in brain cells where the protein does not occur, there is no
physical contact between the $\beta$-globin gene and the LCR.
This paper has been cited about 865 times%
\footnote{\url{http://scholar.google.com/scholar?cites=15030508233271733256}}.
At the time of this publication, de Laat was based at Erasmus University in
Rotterdam, Netherlands but he is now a professor at The Hubrecht Institute in
Utrecht.

Both Dekker \cite{5c} and de Laat \cite{4c} have also worked to develop more
robust versions of the procedure called 5C and 4C respectively.
Each method is designed to improve the flexibility and resolution of the
method by focusing on many smaller fragments of chromatin.
The papers have 412 and 579 citations%
\footnote{\url{http://scholar.google.com/scholar?cites=11459042540325628167}}%
\footnote{\url{http://scholar.google.com/scholar?cites=17950290286653863144}}
respectively.

\subsection*{Part 1.4}

{\bf I chose to focus on 3C for my review} mostly because I found it most
interesting technically.
That being said, it is also an exciting young field that seems, from my
research, to have revolutionized the way that we study the structure of
chromatin.

\newpage

\section*{Part 2}



% === Bibliography ===
\bibliography{midterm}{}
\bibliographystyle{unsrt}

\end{document}
